---
title: "Adele F"
format: html
---

```{r import}
# load data and packages

library(tidyverse)
library(tidytuesdayR)
library(dplyr)
library(forcats)
library(DataExplorer)

tuesdata <- tidytuesdayR::tt_load(2021, week = 29)

scoobydoo <- tuesdata$scoobydoo
```

```{r explore monster numbers}
head(scoobydoo)
dim(scoobydoo)

scoobydoo |>
  group_by(monster_amount) |>
  summarize(number_of_episodes = n())

#Looking deeper into the episodes and observations that have no monsters
scoobydoo |>
  filter(monster_amount == 0) |>
  group_by(series_name, season) |>
  summarize(count = n()) |>
  arrange(desc(count))

scoobydoo |>
  filter(monster_amount == 19)
```

```{r exploring gender ratios}
cat_a <- c("Female")
cat_b <- c("Male")

scoobydoo |>
  group_by(monster_gender) |>
  summarize(number = n()) |>
  filter(monster_gender != "NULL") |>
  mutate(binary_gender = case_when(
    monster_gender %in% cat_a ~ "Female",
    monster_gender %in% cat_b ~ "Male",
    TRUE ~ ">1")) |>
  group_by(binary_gender) |>
  summarize(total = sum(number))

scoobydoo |>
  group_by(culprit_gender) |>
  summarize(number = n()) |>
  filter(culprit_gender != "NULL") |>
  mutate(binary_gender = case_when(
    culprit_gender %in% cat_a ~ "Female",
    culprit_gender %in% cat_b ~ "Male",
    TRUE ~ ">1")) |>
  group_by(binary_gender) |>
  summarize(total = sum(number))


```
From the above table, I learned that 30 episodes have 1 female monster, 319 episodes have 1 male monster, and 167 episodes have more than one monster. Doing the same type of analysis for the number of culprits, I learn tnat there are 116 episodes with more than one culprit, 61 with female culprits, and 263 with male culprits.


```{r explore with DataExplorer package}
scoobydoo |>
  introduce()

plot_intro(scoobydoo)

plot_bar(scoobydoo)

```


```{r explore speech variables}
scoobydoo |>
  filter(jeepers >= 1, jeepers != "NULL") |>
  select(index, jeepers) |>
  arrange(desc(jeepers))

scoobydoo |>
  filter(zoinks >= 1, zoinks != "NULL") |>
  select(index, zoinks) |>
  arrange(desc(zoinks)) |>
  group_by(zoinks) |>
  summarize(n()) |>
  arrange(desc(zoinks))

```

Continuing some gender specific analyses using the data wrangling that Sarah did:
```{r}
# individual monsters of TV series
scoobydoo_clean <- scoobydoo |>
  filter(season %in% c("1", "2","3", "4"),
         format == "TV Series") |>
  filter(monster_name != "NULL") |>
  mutate(monster_name = str_replace(monster_name, "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent,8,Agent,9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal", "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent 8,Agent 9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal")) |>
  pivot_longer(cols = c(monster_name, monster_gender, monster_type, monster_subtype, monster_species), names_to = "monster_var", values_to = "monster") |>
  separate(monster, c("1monster", "2monster", "3monster", "4monster", "5monster", "6monster", "7monster", "8monster", "9monster", "10monster", "11monster", "12monster", "13monster", "14monster", "15monster"), sep = ",") |>
  pivot_longer(cols = ends_with("monster"), names_to = "monster_num", values_to = "monster_value") |>
  pivot_wider(names_from = monster_var, values_from = monster_value) |>
  mutate(monster_num = as.numeric(str_sub(monster_num, 1, 1))) |>
  na.omit()


scoobydoo_monsters <- scoobydoo |>
  filter(season %in% c("1", "2","3", "4"),
         format == "TV Series") |>
  filter(monster_name != "NULL") |>
  mutate(monster_name = str_replace(monster_name, "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent,8,Agent,9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal", "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent 8,Agent 9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal")) |>
  select(title, season, series_name, contains("monster"), contains("unmask")) |>
  pivot_longer(cols = c(monster_name, monster_gender, monster_type, monster_subtype, monster_species), names_to = "monster_var", values_to = "monster") |>
  separate(monster, c("1monster", "2monster", "3monster", "4monster", "5monster", "6monster", "7monster", "8monster", "9monster", "10monster", "11monster", "12monster", "13monster", "14monster", "15monster"), sep = ",") |>
  pivot_longer(cols = ends_with("monster"), names_to = "monster_num", values_to = "monster_value") |>
  pivot_wider(names_from = monster_var, values_from = monster_value) |>
  separate(monster_num, c("monster_num", "x"), sep = "m") |>
  mutate(monster_num = as.numeric(monster_num)) |>
  select(-x) |>
  na.omit() |>
  mutate(monster_type = ifelse(monster_type == "Disugised", "Disguised", monster_type),
          monster_type = ifelse(monster_type == "Disguise", "Disguised", monster_type)) |>
  mutate(monster_type = str_trim(monster_type, side = "left"))


scoobydoo_culprits <- scoobydoo |>
  filter(season %in% c("1", "2","3", "4"),
         format == "TV Series") |>
  filter(culprit_name != "NULL") |>
  select(title, season, series_name, motive, contains("culprit"), zoinks, jeepers, ) |>
  pivot_longer(cols = c(culprit_name, culprit_gender), names_to = "culprit_var",
               values_to = "culprit") |>
  separate(culprit, c("1culprit", "2culprit", "3culprit", "4culprit", "5culprit",
                      "6culprit", "7culprit", "8culprit", "9culprit", "10culprit",
                      "11culprit"), sep = ",") |>
  pivot_longer(cols = ends_with("culprit"), names_to = "culprit_num", values_to = "culprit_value") |>
  pivot_wider(names_from = culprit_var, values_from = culprit_value) |>
  separate(culprit_num, c("culprit_num", "x"), sep = "c") |>
  select(-x) |>
  mutate(culprit_num = as.numeric(culprit_num)) |>
  na.omit()

scoobydoo_clean |>
  filter(monster_type == "Plant")
```
```{r}
scoobydoo_clean |>
  filter(monster_species != "Human") |>
  group_by(monster_gender) |>
  summarize(number = n()) |>
  ggplot(aes(x = monster_gender, y = number)) +
  geom_bar(stat = "identity") +
  labs(title = "Non-Human Monster Genders")

scoobydoo_clean |>
  filter(monster_species == "Human") |>
  group_by(monster_gender) |>
  summarize(number = n()) |>
  ggplot(aes(x = monster_gender, y = number)) +
  geom_bar(stat = "identity")
```

Between human and non-human monsters, the gender ratio seems relatively similar.

```{r gendered motivations exploration}
#This section will explore the relationship between the motivations of culprits and their gender

scoobydoo_culprits |>
  group_by(motive, culprit_gender) |>
  mutate(culprit_gender = str_trim(culprit_gender, side = "left")) |>
  ggplot(aes(x = motive, fill = culprit_gender)) +
  geom_bar(position = "fill", color = "white", linewidth = 0.25) +
  coord_flip() +
  labs(title = "Motives by Gender of Culprit",
       fill = "Culprit Gender",
       x = "Motive",
       y = "Proportion")

```


```{r wrangling}
scoobydoo_monsters |>
  filter(unmask_fred == "TRUE") |>
  group_by(monster_gender) |>
  summarize(n())

scoobydoo_monsters |>
  filter(unmask_daphnie == "TRUE") |>
  group_by(monster_gender) |>
  summarize(n())
  
scoobydoo_monsters |>
  filter(unmask_velma == "TRUE") |>
  group_by(monster_gender) |>
  summarize(n())
  
scoobydoo_monsters |>
  filter(unmask_shaggy == "TRUE") |>
  group_by(monster_gender) |>
  summarize(n())
  
scoobydoo_monsters |>
  filter(unmask_scooby == "TRUE") |>
  group_by(monster_gender) |>
  summarize(n())
  
```







