---
title: "MariaB"
format: html
---

##EDA 
#Guiding question 1: Does the presence of Batman in an episode influence (either) if a trap was used, or the amount of monsters present?
#Guiding question 2: Does a certain type of monster trigger some phrases more than others?


#load packages
```{r}
library(tidyverse)
library(dplyr)
library(DataExplorer) #new package from Michael Clark blog post as shown in slides
```

#load data
```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 29)
scoobydoo <- tuesdata$scoobydoo
```


```{r}
head(scoobydoo)
```

```{r}
str(scoobydoo)
```

#Report using data explorer package
```{r}
create_report(scoobydoo)

```
#phrases that are said multiple times (more than 2) in an episode: 'zoinks', 'jinkies'
#phrases said mostly once or twice: 'sooby doo where are you', 'rooby rooby roo', 'my glasses', 'just about wrapped up' 
#133 monster categories



##For guiding question 1: Does the presence of Batman in an episode influence (either) if a trap was used, or the amount of monsters present?



```{r}
# Convert to factor if needed
#data$batman_present <- as.factor(data$batman_present)
scoobydoo$set_a_trap <- as.factor(scoobydoo$set_a_trap)

# Check proportions
table(scoobydoo$batman, scoobydoo$set_a_trap)
prop.table(table(scoobydoo$batman, scoobydoo$set_a_trap), 1)



```
#On average, there are more monsters present when batman is in an episode. But, there are many large outliers when batman is not present in an episode. 
```{r}
ggplot(scoobydoo, aes(x = batman, y = monster_amount)) +
  geom_boxplot() 

```





##For guiding question 2: Does a certain type of monster trigger some phrases more than others?

#how often are the phrases said?
```{r}
scoobydoo %>%
  count(zoinks)
#26 times in one episode

scoobydoo %>%
  count(groovy)
#14 times in one episode

scoobydoo %>%
  count(jeepers)
#10 times in one episode

scoobydoo %>%
  count(rooby_rooby_roo)
#7 times in one epsiode


scoobydoo %>%
  count(scooby_doo_where_are_you)
#4 times in one episdoe


```

#What type of monsters are there and how often do they come up in the data?
```{r}
scoobydoo |> 
  count(monster_type)
```

```{r}
scoobydoo %>%
  count(monster_type, sort = TRUE) %>%
  ggplot(aes(x = reorder(monster_type, -n), y = n, fill = monster_type)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Count of Monster Types", x = "Monster Type", y = "Count")

```

##NEW EDA with wrangled data (monsters separated) AND new question
#First wrangle data with code from Sarah

```{r}
# individual monsters of TV series
scoobydoo_clean <- scoobydoo |>
  filter(season %in% c("1", "2","3", "4"),
         format == "TV Series") |>
  filter(monster_name != "NULL") |>
  mutate(monster_name = str_replace(monster_name, "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent,8,Agent,9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal", "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent 8,Agent 9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal")) |>
  select(title, season, series_name, contains("monster")) |>
  pivot_longer(cols = c(monster_name, monster_gender, monster_type, monster_subtype, monster_species), names_to = "monster_var", values_to = "monster") |>
  separate(monster, c("1monster", "2monster", "3monster", "4monster", "5monster", "6monster", "7monster", "8monster", "9monster", "10monster", "11monster", "12monster", "13monster", "14monster", "15monster"), sep = ",") |>
  pivot_longer(cols = ends_with("monster"), names_to = "monster_num", values_to = "monster_value") |>
  pivot_wider(names_from = monster_var, values_from = monster_value) |>
  mutate(monster_num = as.numeric(str_sub(monster_num, 1, 1))) |>
  na.omit()
```

#New question:
#Does the gender of a character influence their likelihood of being portrayed as a monster or criminal, and are there specific phrases that consistently appear in episodes featuring female vs. male villains?


#Gender of monsters
```{r}
ggplot(scoobydoo_clean, aes(y = monster_type, fill = monster_gender,)) + 
  geom_bar(position = "fill")
```











