---
title: "Report"
execute: 
  echo: false      # change to true to show the code
  warning: false
code-fold: false  # change to true to fold the code chunks
  
---


```{r}
# load data and packages

library(tidyverse)
library(tidytuesdayR)
library(patchwork)
library(gt)

tuesdata <- tidytuesdayR::tt_load(2021, week = 29)

scoobydoo <- tuesdata$scoobydoo
```

```{r}
# wrangling data


# wrangle/clean monster data 
scoobydoo_monsters <- scoobydoo |>
  filter(season %in% c("1", "2","3", "4"),
         format == "TV Series") |>
  filter(monster_name != "NULL") |>
  mutate(unmasked = case_when(
    unmask_fred == TRUE | unmask_daphnie == TRUE | unmask_velma == TRUE | 
    unmask_shaggy == TRUE | unmask_scooby == TRUE | unmask_other == TRUE ~ "TRUE", 
    TRUE ~ "FALSE")) |>
  mutate(monster_amount = ifelse(title == "The Ghastly Ghost Town", 5, monster_amount),
         monster_amount = ifelse(title == "Me and My Shadow Demon", 3, monster_amount)) |>
  mutate(monster_name = str_replace(monster_name, "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent,8,Agent,9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal", "Dr. Phineus Phibes,Agent 1,Agent 2,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent 8,Agent 9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal"),
         monster_name = str_replace(monster_name, "Dr. Phineus Phibes,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent,9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal,Misty Swiss", "Dr. Phineus Phibes,Agent 3,Agent 4,Agent 5,Agent 6,Agent 7,Agent 9,Agent 10,Agent 11,Agent 12,Agent 13,Dr. Trebal,Misty Swiss"),
         monster_name = str_replace(monster_name, "Dr. Phineus Phibes,Dr. Trebla,Agent 1,Agent 2,Mark,Ricky,Agent,4,Agent 7,Agent 11", "Dr. Phineus Phibes,Dr. Trebla,Agent 1,Agent 2,Mark,Ricky,Agent 4,Agent 7,Agent 11")) |>
  select(title, season, series_name, contains("monster"), unmasked) |>
  pivot_longer(cols = c(monster_name, monster_gender, monster_type, monster_subtype, monster_species), names_to = "monster_var", values_to = "monster") |>
  separate(monster, c("1monster", "2monster", "3monster", "4monster", "5monster", "6monster", "7monster", "8monster", "9monster", "10monster", "11monster", "12monster", "13monster", "14monster", "15monster"), sep = ",") |>
  pivot_longer(cols = ends_with("monster"), names_to = "monster_num", values_to = "monster_value") |>
  pivot_wider(names_from = monster_var, values_from = monster_value) |>
  separate(monster_num, c("monster_num", "x"), sep = "m") |>
  mutate(monster_num = as.numeric(monster_num)) |>
  select(-x) |>
  na.omit() |>
  mutate(monster_type = ifelse(monster_type == "Disugised", "Disguised", monster_type),
         monster_type = ifelse(monster_type == "Disguise", "Disguised", monster_type),
         monster_type = ifelse(monster_type == "Dr. Trebal", "Super-Villain", monster_type)) |>
  mutate(monster_name = ifelse(monster_name == "Dr Trebla", "Dr. Trebla", monster_name),
         monster_name = ifelse(monster_name == "Dr. Trebal", "Dr. Trebla", monster_name)) |>
  mutate(monster_type = str_trim(monster_type, side = "left"))



# wrangle/clean culprit data
scoobydoo_culprits <- scoobydoo |>
  filter(season %in% c("1", "2","3", "4"),
         format == "TV Series") |>
  filter(culprit_name != "NULL") |>
  select(title, season, series_name, motive, contains("culprit")) |>
  pivot_longer(cols = c(culprit_name, culprit_gender), names_to = "culprit_var",
               values_to = "culprit") |>
  separate(culprit, c("1culprit", "2culprit", "3culprit", "4culprit", "5culprit",
                      "6culprit", "7culprit", "8culprit", "9culprit", "10culprit",
                      "11culprit"), sep = ",") |>
  pivot_longer(cols = ends_with("culprit"), names_to = "culprit_num", values_to = "culprit_value") |>
  pivot_wider(names_from = culprit_var, values_from = culprit_value) |>
  separate(culprit_num, c("culprit_num", "x"), sep = "c") |>
  select(-x) |>
  mutate(culprit_num = as.numeric(culprit_num)) |>
  na.omit() |>
  mutate(culprit_gender = str_trim(culprit_gender, side = "left"))
```
# Motivation
We chose to work with the Scooby Doo dataset due to the wide range of variables that were addressed across all episodes of Scooby Doo released. There is a lot of potential for the incorporation of fun visuals into our final visualizations, as our dataset comes from a familiar entertainment show that people can relate to. It will be fun to statistically analyze Scooby Doo and shine new light on something that many people were likely familiar with in their childhoods. Particularly, looking at the data set initially, we were intrigued by the prospect of correlating certain factors (certain phrases said, number of monsters, voice actors involved) to each other across the broad range of data collected for each episode. This project is important because it may reveal patterns that many people watching TV were exposed to when they were younger (gender bias in who is a monster and criminal? Which phrases and terms are associated with which scary events?) that they should consider when thinking about bias and actions in their lives now.


# Research Question
* Does the gender of a character in Scooby Doo episodes influence their likelihood of being portrayed as a monster or criminal?
* Are females or males unfairly represented across Scooby Doo episodes?



# Background
We are working with the Scooby Doo dataset from Kaggle, and we collected a set of phrases from all episodes of the Scooby Doo series. We obtained this data from plummye’s dataset, where the Kaggle user watched every episode of Scooby Doo and recorded the variables over the course of a year. They collected data for 75 variables about the title of the episode, voice actors in the episode, who the monsters and villains were, and what key phrases were used in the episode. The data set includes Scooby Doo movies and specials as well, and which phrases were focused on are those which are known to be common in and unique to Scooby Doo, such as “zoinks,” “jeepers,” and “scooby snacks.”



# Data
The dataset contains information on different variables for every episode or movie of Scooby-Doo. According to the creator of the dataset, user Plummye on Kaggle, they made the dataset because they struggled to find a dataset that was “interesting, large, unstudied, [and] many different types of variables.” So, Plummye decided to make their own dataset for Scooby-Doo episodes because they were watching a lot of Scooby-Doo during the 2020 lockdown. It took them over a year to watch all the episodes and movies, plus tabulate all the data. We found the dataset on TidyTuesday (week 07/13/2021), but it was originally posted on Kaggle by user Plummye. 

There are many variables tabulated for each episode. In total, there are 75 variables, and data is available for 603 Scooby-Doo episodes and movies. The types of measurements vary from variable to variable, with some being numerical (ie IMDB score, runtime, number of monsters), some are categorical (ie episode name, type of monster, voice actor names), and some are boolean (ie if a certain line/catchphrase was said). Some entries for variables are “NULL,” so we will have to filter them out when working with the data. Additionally, when there is more than one monster, they are listed in one row, separated by commas (ie monster a,monster b,monster c), so this needs to be reformatted as well.



# Data Insights
## Exploration of Monster Genders
```{r}
# graph of male vs female monsters
```


```{r}
#| fig-alt: "Bar graph showing the percentage of monster genders (male, female, and none) for each Scooby-Doo TV series."

# monster genders (%) by series
scoobydoo_monsters |>
  group_by(monster_gender) |>
  mutate(n = n()) |>
  mutate(monster_gender = as.factor(monster_gender)) |>
  mutate(monster_gender = fct_relevel(monster_gender, c("Female", "None", "Male"))) |>
  ggplot(aes(y = series_name, fill = monster_gender)) +
  geom_bar(position = "fill")+
  labs(x = "Percent", y = "Series Name",
       title = "Percentages of Monsters by Gender of TV Series",
       fill = "Monster Gender",
       caption = "Data collected by @plummye on Kaggle, used for week 07/13/2021 of Tidy Tuesday")+
  scale_fill_viridis_d()+
  theme(plot.caption = element_text(face = "italic"),
        plot.title = element_text(face = "bold"))
```
:::{.callout-note title="Explanation"}
write expalantion here
:::

```{r}

#| fig-alt: "This is a bar chart of monster type by gender. On the y-axis, there are all the monster types. On the x-axis, are decimals counting up to 1. The bars show the proportion of each monster that is male, female, and none. The monster type with most female versions is magician. More than half magicians are female. After that, not many monsters have female versions – undead, ancient, mythical, mechanical have a few female versions, but less than a quarter of those monsters are female. "

scoobydoo_monsters |>
  mutate(monster_gender = as.factor(monster_gender)) |>
  mutate(monster_gender = fct_relevel(monster_gender, "Female", "None", "Male")) |>
  ggplot(aes(y = monster_type, fill = monster_gender)) + 
  geom_bar(position = "fill")+
    expand_limits(x = 1.25)+
  coord_cartesian(xlim = c(0,1.3))+
  geom_rect(aes(xmin = -0, xmax = 1, ymin = 5.49, ymax = 6.51),
            color = "red", fill = NA, linewidth = .75)+
  annotate("text", x = 1.17, y = 6, label = "Most magicians \nare female?", 
           size = 3.2)+
  annotate("text", x = 1.18, y = 9, label = "*The monster w/ no \ngender is a plant",
           size = 3, fontface = "italic")+
  labs(x = "Proportion of Each Gender", y = "Monster Type", 
       title = "Breakdown of Monster Types by Gender",  fill = "Monster Gender",
       subtitle = "Gender determined by the pronouns used to refer to the monster",
       caption = "Data collected by @plummye on Kaggle, used for week 07/13/2021 of Tidy Tuesday") +
  scale_fill_viridis_d() +
  theme(plot.caption = element_text(face = "italic"),
        plot.title = element_text(face = "bold", size = 14))
  
```

:::{.callout-note title="Explanation"}
write expalantion here
:::



```{r}
#| fig-alt: "Bar graph showing the proportion of monster types for each Scooby-Doo TV series"

# monster type by series
scoobydoo_monsters |>
  ggplot(aes(y=series_name, fill = monster_type))+
  geom_bar(position = "fill", color = "white")+
  scale_fill_viridis_d()+
  labs(title = "Proportion of Monster Types by Series",
       x = "Proportion", y = "Series Name", fill = "Monster Type",
       caption = "Data collected by @plummye on Kaggle, Week 07/13/2021 of Tidy Tuesday")+
  theme(plot.title = element_text(face = "bold", size = 14, hjust=1),
        plot.caption = element_text(face = "italic"))
```
:::{.callout-note title="Explanation"}
For the series "Shaggy & Scooby-Doo Get a Clue," a the proprtion of monster types varies a lot amongst other series. It has the largest proportion of super-villains and disguised monsters, but very few ghosts. For other series, ghost and undead are the most common monster types.
:::

## Female Monsters over Time
```{r}
# code/graph tba
```

:::{.callout-note title="Explanation"}
write expalantion here
:::

## Portrayal of Male vs Female Monsters
```{r}
#| fig-alt: "Graphs comparing magician monster subtypes by gender. The first graph shows most female magician monsters are witches, while male magicians are wizards. The second graph shows for all witch subtypes, only 1 is male. For all wizard subtypes, none are female."

# female monsters are most commonly portrayed as magicians, specifically witches

p1 <- scoobydoo_monsters |> 
  filter(monster_type == "Magician") |>
  count(monster_subtype, monster_gender) |>
  ggplot(aes(y=n,x=monster_subtype, fill = monster_gender))+
  geom_bar(stat = "identity")+
  labs(title = "Subtypes of Magician Monsters, by Gender",
       y = "Count", x = "Monster Subtype",
       fill = "Monster Gender") +  
       scale_fill_viridis_d()+
  scale_y_continuous(breaks = seq(0, 10,by=2))+
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- scoobydoo_monsters |>
  filter(monster_subtype %in% c("Witch", "Wizard")) |>
  ggplot(aes(x=monster_subtype, fill = monster_type))+
  geom_bar()+
  labs(title = "Witch and Wizard Monsters by Gender",
       x = "Monster Subtype", y = "Number of Monsters",
       fill = "Monster Type",
       caption = "Data collected by @plummye on Kaggle, used for week 07/13/2021 of Tidy Tuesday")+
  facet_wrap(~monster_gender) +
  scale_fill_viridis_d()+
  theme(plot.caption = element_text(face = "italic"),
        plot.title = element_text(face = "bold", size = 14))+
    scale_y_continuous(breaks = seq(0, 10,by=2))


p1 / p2
```
:::{.callout-note title="Explanation"}
write explanation here
:::


```{r, fig.width=8}
#| fig-alt: "2 bar graphs exploring if a monster is real (true or false) and if a monster is unmasked (true or false) varies by gender. A larger percentage of male monsters are real and not unmasked, compared to female monsters."
# meanwhile, male monsters are more recurring than female monsters and are treated like antagonists

p1 <- scoobydoo_monsters |> 
  filter(monster_gender != "None") |>
  ggplot(aes(x=monster_gender, fill = monster_real))+
  geom_bar(position = "fill")+
  labs(title = "'Monster Real' Status by Gender",
       subtitle = "Is the monster an actual supernatural creature (T), \nor just someone pretending to be one (F)?",
       x = "Monster Gender", y = "Percentage of Monsters", fill = "Real?")+
  scale_fill_viridis_d()+
  theme(plot.caption = element_text(face = "italic"),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
        legend.position = "bottom")+
  scale_y_continuous(labels = scales::percent_format())

p2 <- scoobydoo_monsters |> 
  filter(monster_gender != "None") |>
  ggplot(aes(x = monster_gender, fill=unmasked))+
  geom_bar(position = "fill")+
  labs(title = "'Unmasked' Status by Gender",
       subtitle = "Was the monster revealed to be a disguised \nculprit (T) by the Gang during the episode or not (F)?",
       x = "Monster Gender", y = "Percentage of Monsters", fill = "Unmasked?",
       caption = "Data collected by @plummye on Kaggle, used for week 07/13/2021 of Tidy Tuesday")+
  scale_fill_viridis_d()+
  theme(plot.caption = element_text(face = "italic"),
        plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        legend.position = "bottom")+
  scale_y_continuous(position = "right",
                     labels = scales::percent_format())

p1 | p2

# male monsters are more likely to be portrayed as real (ie not a culprit in disguise) and are less likely to be unmasked
```
:::{.callout-note title="Explanation"}
Within Scooby-Doo episodes, the monster can either be real or not real. Most commonly, a monster is not real, meaning it is just a person in disguise (ie the culprit) and the monster's role was to scare the audience and the Gang. These monsters are commonly unmasked by the Gang in the show, which was tracked by the variable "unmasked." However, there are instances when the monster is actually real and not a person in disguise. In these instances, the monster is often not unmasked, or the Gang unmaskes a version of that monster in disguise.  


From the graph on the left, there is a higher proportion of male monsters that are real (~30%) compared to real female monsters (~20%). There is also a higher proportion of female monsters unmasked (over 50% or half), while only a third of male monsters are unmasked, as indicated by the chart on the right. From these two visualizations, there is a clear trend that when there are female monsters in episodes, they are typically portrayed as an ordinary monster in disguise that is unmasked by the Gang. While there are some male monsters portrayed like this as well, there are a greater number of male monsters that are real and are never unmasked. As a follow-up, we will look at more into male monsters that are not unmasked/real monsters to see if there are any trends for this kind of monster in the show.   

:::

```{r}
#| fig-alt: "Bar graph showing the most popular types of male monsters based on if the monster is real (true or false) and unmasked (true or false). Most non-real monsters are animals or ghosts, while real monsters are commonly disguised or super-villains."
# look at non-unmasked or real male monsters

# labels
label_positions <- tibble(
  real_unmasked = c("TRUE, FALSE", "TRUE, FALSE", "FALSE, TRUE", "FALSE, TRUE", 
                    "FALSE, FALSE", "FALSE, FALSE"),
  monster_type = c("Super-Villain", "Disguised", "Animal", "Ghost", "Animal", "Ghost"),
  y_position = c(50, 200, 220, 140, 145, 85))


# plot
scoobydoo_monsters |>
  filter(monster_gender == "Male") |>
  mutate(real_unmasked = paste(monster_real, unmasked, sep= ", ")) |>
  ggplot(aes(fill=monster_type, x = real_unmasked))+
  geom_bar(position = "stack", color = "white")+
  geom_label(data = label_positions, 
            aes(x=real_unmasked, label = monster_type, y = y_position),
            color = "black", fill = "white")+
  labs(title = "Types of Male Monsters based on Monster Real and Unmasked",
       x = "Monster Real (T/F), Unmasked (T/F)", y = "Number of Monsters",
       fill = "Monster Type",
       caption = "Data collected by @plummye on Kaggle, used for week 07/13/2021 of Tidy Tuesday")+
  scale_fill_viridis_d()+
  theme(plot.caption = element_text(face = "italic"))
  




```

:::{.callout-note title="Explanation"}
Ghosts and animals are the most common type of monsters that are not real, and also tend to be unmasked a decent amount of times. Real monsters are commonly disguised, but there are also a large number of super-villains. There are very few monsters that are real and unmasked.

From an earlier graph, one of main monster types for the series "Shaggy & Scooby-Doo Get a Clue!" was super-villain. We are curious if the series has a lot of super-villains, or if the super-villain monster is recurring because it is real and the Gang tries to unmask it for the duration of the series. Or maybe because it is a super-villain, it has a long-term plan to sabotage the Gang, hence the large number of apperances. 
:::

```{r}
# "Shaggy and Scooby-Doo Get a Clue!" has a lot of super-villains and disguised monsters, super-villain monsters in the show are almost entirely male, and super-villains + disguised monsters are commonly real monsters, which are not unmasked. What's up?


scoobydoo_monsters |>
  filter(series_name == "Shaggy & Scooby-Doo Get a Clue!") |>
  count(monster_name, monster_type, monster_gender, monster_real, unmasked) |>
  arrange(desc(n)) |>
  gt() |>
  tab_header(title = "Monsters in 'Shaggy & Scooby-Doo Get a Clue!'") |>
  cols_label(monster_name = "Name", monster_type = "Type", monster_gender = "Gender",
             monster_real = "Real", unmasked = "Unmasked", 
             n = "Times Seen") |>
  cols_align(align = "center")

# ok so Dr. Phineus Phibes is a super-villain from the series "Shaggy..." and because he appears in probably all the episodes, is real, and not unmasked, he's treated like the main antagonist of the show (as opposed to the typical masked monster/unmasked culprit format). 
# also Dr. Trebla (another super-villain from same series)
# also why must he have 13 agents...that are all male?
# 
```
:::{.callout-note title="Explanation"}
According to the table, most of the monsters from the series "Shaggy & Scooby-Doo Get a Clue" are male, real, and not unmasked. The individual monsters (based on monster_name), seem to appear in multiple episodes from the series. For instance, Dr. Phineus Phibes appears in 25 episodes and Dr. Trebla appears in 23. This likely indicates these super-villains are recurring monsters or antagonists for the this series, with long-term plans to sabotage Scooby-Doo and the Gang, hence why they are real and not unmasked. 

However, there are no female super-villains, or monsters that appear this many times and are the main villain for an entire series. This is consistent with how male villains are commonly portrayed in movies and TV. Male villains are commonly depicted as strategic masterminds, scientists, and inventors that use technology or calculated plans to achieve world dominance, which is similar with how Dr. Phineus Phibes is portrayed in the show. The monsters in the show don't follow the typical structure of one-off culprits, but have long-term schemes with greater character development, showcasing a possible lack of investment in female antagonists compared to male antagonists.

This series was also released in 2006-2008, so it is relatively recent, especially compared to the older episodes of Scooby-Doo.

:::

## Culprit Motives
```{r}
# code/graph tba
```

:::{.callout-note title="Explanation"}
write explantion here
:::




# Conclusions
* Scooby Doo episodes favor representation of male villains and culprits
* Females are restricted to certain roles and titles (often based on gender stereotypes)
* Little change in representation over time (but lesser appearances over time)


# Limitations and Future Work
* Data was collected by one person – subject to error
* Subjective variable definitions


* Are there specific phrases that consistently appear in episodes featuring female vs. male villains?
* Are there similar patterns of gender stereotypes with other monster types or in different series?
* How do these findings translate to Scooby-Doo movies? Are male and female monsters portrayed similarly in movies? 
* Does the portrayal of female monsters change over time? Do they have less stereotypical roles in more recent episodes?
